# Advanced Topics

In this chapter, we talk about a few advanced topics that may be of interest to developers and advanced users.

## More global options

There are a few more advanced global options in addition to those introduced in Section \@ref(global-options), and they are listed in Table \@ref(tab:global-options2).

```{r global-options2, echo=FALSE}
knitr::kable(matrix(c(
  'blogdown.hugo.dir', '', 'the directory of the Hugo executable',
  'blogdown.method', 'html', 'the building method for R Markdown',
  'blogdown.publishDir', '', 'the publish dir for local preview',
  'blogdown.widgetsID', TRUE, "incremental ID's for HTML widgets?",
  NULL
), ncol = 3, byrow = TRUE, dimnames = list(NULL, c('Option name', 'Default', 'Meaning'))), booktabs = TRUE, caption = 'A few more advanced global options.')
```

If you want to install Hugo to a custom path, you can set the global option `blogdown.hugo.dir` to a directory to store the Hugo executable before you call `install_hugo()`, e.g., `options(blogdown.hugo.dir = '~/Downloads/hugo_0.20.1/')`. This may be useful for you to use a specific version of Hugo for a specific website,^[You can set this option per project. See Section \@ref(global-options) for details.] or store a copy of Hugo on a USB Flash drive along with your website.

The option `blogdown.method` is explained in Section \@ref(methods).

When your website project is under version control in the RStudio IDE, it may be slow to preview the site continuously if it contains hundreds of files or more, because the default publish directory is `public/` under the project root directory, and whenever you make a change in the source that triggers a rebuild, RStudio will be busy with tracking file changes in the `public/` directory. The delay before you see the website in the RStudio Viewer can be more than 10 seconds or even longer. That is why we provide the option `blogdown.publishDir`. You may set a temporary publish directory to generate the website, and this directory should not be under the same RStudio project, e.g., `options(blogdown.publishDir = '../public_site')`, which means the website will be generated to the directory `public_site/` under the parent directory of the current project.

The option `blogdown.widgetsID` is only relevant if your website source is under version control and you have HTML widgets on the website. If this option is `TRUE` (default), the random ID's of HTML widgets will be changed to incremental ID's in the HTML output, so that these ID's are unlikely to change every time you recompile your website, otherwise every time you will get different random ID's.

## Different building methods {#methods}

If your website does not contain any Rmd files, it is very straightforward to render it --- just a system call to the `hugo` command. When your website contains Rmd files, **blogdown** has provided three different rendering methods to compile these Rmd files. A website can be built using the function `blogdown::build_site()`:

```{r eval=FALSE, code=formatR::usage(blogdown::build_site, output=FALSE), tidy=FALSE}
```

As mentioned in Section \@ref(global-options), the default value of the `method` argument is determined by the global option `blogdown.method`, and you can set this option in `.Rprofile`.

Both `html` and `html_encoded` use `rmarkdown::render()` to render Rmd files to HTML, and the technical differences between them are:

1. `method = 'html'` renders `*.Rmd` to `*.html`, and keeps the `*.html` output file under the same directory as `*.Rmd`; `method = 'html_encoded'` will move `*.html` to the directory `/blogdown/`.

1. An Rmd file may generate two directories for figures (`*_files/`) and cache (`*_cache/`), respectively, if you have plot output or HTML widgets [@R-htmlwidgets] in your R code chunks, or enabled the chunk option `cache = TRUE` for caching. In the figure directory, there will be a subdirectory `figure-html/` that contains your plot output files, and possibly other subdirectories containing HTML dependencies from HTML widgets (e.g., `jquery/`). For `method = 'html'`, it moves the figure directory to `/static/`, and the cache directory to `/blogdown/`. For `method = 'html_encoded'`, all output files and directories are moved to `/blogdown/`, including figures, cache, and `*.html`.

1. The name of the method `'html_encoded'` comes from the fact that the paths of R plot files and HTML widgets are first encoded in the `*.html` output, then Hugo is called to build the website, and finally all HTML files under the `/public/` directory are post-processed to decode the paths. The main reason for this trouble is that it is not easy to predict where a certain page will be rendered in the `/public/` directory,^[For example, a source file `/content/about.md` may be rendered to `/public/about/index.html`, and `/content/post/2017-02-14-hello.Rmd` could be rendered to `/public/2017/02/14/hello/index.html` (the content of the post could also be used in `/public/categories/index.html`). It dependes on the rules for permanent links you set for Hugo.] so the relative paths of figures and widgets may fail to work after a site is rendered, and they need to be resolved by post-processing. By comparison, `method = 'html'` does not need post-processing due to the fact that the relative paths will be resolved to absolute paths at the pre-processing stage.

We must have successfully confused you now, but do not worry. You do not have to understand these technical details, unless you want to help us with hacking. For an average user, you only need to know the pros and cons of the two methods.

- For `method = 'html'`:

    - Your `/content/` directory may not be completely clean, since it also contains `*.html` files.
    
    - If you use HTML widgets in Rmd, the HTML pages may not conform to W3C standards,^[W3C stands for World Wide Web Consortium: https://www.w3.org. In case you are really curious: the problem with `method = 'html'` for HTML widgets is that CSS styles may be declared in the `<body>` tag.] but the widgets will be most likely to work well anyway.

    - Besides these relatively minor disadvantages, the major advantage of `method = 'html'` is that your website is ready to be compiled by Hugo _alone_ after you have compiled it through **blogdown** once, because there is no need to use R or **blogdown** to post-process the website like `method = 'html_encoded'`. This advantage gives you the freedom to use deploying services like Netlify (Chapter \@ref(deployment)), where neither R nor **blogdown** is not available, but Hugo is. It is also much faster than `'html_encoded'.`

- For `method = 'html_encoded'`:

    - Your `/content/` directory is free of output files.
    
    - The HTML files under `/public/` should conform to standards.
    
    - The disadvantages are: rendering the website is dependent on **blogdown** all the time, and it is relatively slow.^[It should not be too bad. Typically a website can be rendered in a few seconds.]

No matter which method you choose to use, `build_site()` will always look for an R script `/R/build.R` and execute it if it exists. This gives you the complete freedom to do anything you want for the website. For example, you can call `knitr::knit()` to compile Rmd to Markdown (`*.md`) in this R script instead of using `rmarkdown::render()`. This feature is designed for advanced users who are really familiar with the **knitr** package^[Honestly, it was originally designed for Yihui himself to build his own website, but he realized this feature could actually free users from Hugo. For example, it is possible to use Jekyll (another popular static site generator) with **blogdown**, too.] and Hugo or other static website generators (see Chapter \@ref(other-generators)).

For `method = 'custom'`, you are completely on your own and should be responsible for building the website using `/R/build.R`; `build_site()` will not call `rmarkdown::render()` or Hugo at all in this case.

The argument `local` of `build_site()` indicates whether you want to build the website for local preview or publishing. In the former mode (`local = TRUE`), a few Hugo options are temporarily changed when building the site:

- `baseurl` is set to `/`, which means the root URL is `/`, even if you have set it to another value like `http://example.com/` in `config.toml`. This is because we do not really want the URLs in our website to have the actual base URL when previewing the website locally (e.g., you may be offline and cannot connect to `http://example.com/`).

- Posts and pages with the options `draft: true` and/or `future: true` in the YAML metadata are always rendered, i.e., drafts and posts with future publication dates are rendered in the local preview mode.

In the latter mode (`local = FALSE`), no options are modified, and all your options set in `config.toml` and YAML metadata are respected.

The function `blogdown::serve_site()` calls `blogdown::build_site(local = TRUE)`. Even if the website generated by `serve_site()` under `/public/` is a functional website, we still recommend that you call `blogdown::build_site(local = FALSE)`^[You can omit `local = FALSE` since it is the default.] before you publish the `/public/` folder. If your drafts and future posts contain (time-)sensitive information, you are strongly recommended to delete the `/public/` directory before you rebuild the site for publishing every time, because Hugo never deletes it, and your sensitive information may be rendered by a certain `build_site(local = TRUE)` call last time and left in the directory. If the website is really important, and you need to make sure you absolutely will not screw up anything every time you publish it, put the `/public/` directory under version control, so you have a chance to see which files were changed before you publish the new site.

## Building a website for local preview {#local-preview}

The function `blogdown::build_site()` has an argument `local` that defaults to `FALSE`, which means building the website for publishing instead of local previewing. The mode `local = TRUE` is primarily for `blogdown::serve_site()` to serve the website locally. There are two major differences between `local = FALSE` and `TRUE`. When `local = TRUE`:

- The `baseurl` option in `config.toml` is temporarily overridden by `"/"` even if you have set it to a full URL like `"http://www.example.com/"`.^[If your `baseurl` contains a subdirectory, it will be overridden by the subdirectory name. For example, for `baseurl = "http://www.example.com/project/"`, `build_site(local = TRUE)` will temporarily remove the domain name and only use the value `/project/`.] This is because when a website is to be previewed locally, links should refer to local files. For example, `/about/index.html` should be used instead of the full link `http://www.example.com/about/index.html`; `serve_site()` knows that `/about/index.html` means the file under the `public/` directory, and can fetch it and display the conten to you, otherwise your browser will take you to the website `http://www.example.com` instead of displaying a local file.

- Draft and future posts are always rendered when `local = TRUE`, but not when `local = FALSE`. This is for you to preview draft and future posts locally. If you know the [Hugo command line](https://gohugo.io/commands/hugo/), it means the `hugo` command is called with the flags `-D -F`, or equivalently, `--buildDrafts --buildFuture`.

- There is a caching mechanism to speed up building your website: an Rmd file will not be recompiled when its `*.html` output file is newer (in terms of file modification times). If you want to force `build_site(local = TRUE)` to recompile the Rmd file even if it is older than the HTML output, you need to delete the HTML output, or edit the Rmd file so that its modification time will be newer.

If your website does not work without the full `baseurl`, or you do not want the draft or future posts to be published, you should not publish the `public/` directory generated by `serve_site()`. Always run `blogdown::build_site()` before you upload this directory to a web server. You do not need to worry about these details if your website is automatically generated from source via a service like Netlify, which will make use of `baseurl` and not use `-D -F` by default.

## The default HTML template {#default-template}

As we mentioned in Section \@ref(output-format), the default output format for an Rmd document in **blogdown** is `blogdown::html_page`. This format passes a minimal HTML template to Pandoc by default:

```{r default-template, engine='cat', code=readLines(blogdown:::pkg_file('resources', 'template-minimal.html')), class.source='html'}
```

You can find this template file via `blogdown:::pkg_file('resources', 'template-minimal.html')` in R, and this file path is the default value of the `template` argument of `html_page()`. You may change this default template, but you should understand what this template is supposed to do first.

If you are familiar with Pandoc templates, you should realize that this is not a complete HTML template, e.g., it does not have the tags `<html>`, `<head>`, or `<body>`. That is because we do not need or want Pandoc to return a full HTML document to us. The main thing we want Pandoc to do is to convert our Markdown document to HTML, and give us the body of the HTML document, which is in the template variable `$body$`. Once we have the body, we can further pass it to Hugo, and Hugo will use its own template to embed the body and generate the full HTML document. Let's explain this by a minimal example. Suppose we have an R Markdown document `foo.Rmd`:

```markdown
---
title: "Hello World"
author: "Yihui Xie"
---

I found a package named **blogdown**.
```

It is first converted to an HTML file `foo.html` through `html_page()`, and note that YAML metadata are ignored for now:

```html
<!-- BLOGDOWN-HEAD -->
<!-- /BLOGDOWN-HEAD -->

I found a package named <strong>blogdown</strong>.
```

Then **blogdown** will read the YAML metadata of the Rmd source file, and insert the metadata into the HTML file so it becomes:

```html
---
title: "Hello World"
author: "Yihui Xie"
---

I found a package named <strong>blogdown</strong>.
```

This is the file to be picked up by Hugo and eventually converted to an HTML page of a website. Since the Markdown body has been processed to HTML by Pandoc, Hugo will basically use the HTML. That is how we bypass Hugo's Markdown engine BlackFriday.

Besides `$body$`, you may have noticed other Pandoc template variables like `$header-includes$`, `$css$`, `$include-before$`, `$toc$`, and `$include-after$`. These variables make it possible to customize the `html_page` format. For example, if you want to generate a table of contents, and apply an additional CSS stylesheet to a certain page, you may set `toc` to `true` and pass the stylesheet path to the `css` argument of `html_page()`, e.g.,

```yaml
---
title: "Hello World"
author: "Yihui Xie"
output:
  blogdown::html_page:
    toc: true
    css: "/css/my-style.css"
---
```

There is also a pair of HTML comments in the template: `<!-- BLOGDOWN-HEAD -->` and `<!-- /BLOGDOWN-HEAD -->`. This is mainly for `method = 'html_encoded'` in `blogdown::build_site()` (see Section \@ref(methods)). This pair of comments are used to mark the HTML code fragment that should be moved to the `<head>` tag of the final HTML page. Typically this code fragment contains links to CSS and JavaScript files, e.g., those requested by the user via the `css` argument of `html_page()`, or automatically generated when HTML widgets are used in an Rmd document. For `method = 'html'`, this code fragment is not moved, which is why the final HTML page may not conform to W3C standards. If you want to customize the template, you are recommended to use this pair of comments to mark the HTML code fragment that belongs to the `<head>` tag.
